---
title: "HW1: An Exploration of the Gapminder DataSet, and My First R Markdown Experience"
author: "VIcki Hertzberg"
date: "January 18, 2017"
output: html_document
---

The link to this file is as follows: 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gapminder)
```

## Gapminder Example

We are working with the gapminder dataset. Here is a summary of the data.

```{r gapminder}
summary(gapminder)
```

There are 1704 observatons with 6 variables. The names of the variables are `r names(gapminder)`.

Oh cool, I made that last sentence with an inline r command. Neat, huh!

Here is the structure of the dataset:

```{r}
str(gapminder)
```

The mean life expectancy is `r round(mean(gapminder$lifeExp), digits = 2)` and the standard deviation is `r round(sd(gapminder$lifeExp), digits = 2)`. Median life expectancy is `r round(median(gapminder$lifeExp), digits = 2)` and the number of observations is `r round(length(gapminder$lifeExp), digits = 2)`.

In the little table below we see the median life expectancy by continent:

```{r}
aggregate(lifeExp ~ continent, gapminder, median)
```

Here we see the median life expectancy by year:

```{r}
aggregate(lifeExp ~ year, gapminder, median)
```


Here is the mean life expectancy by continent:

```{r}
aggregate(lifeExp ~ continent, gapminder, mean)
```

and the standard deviation:

```{r}
aggregate(lifeExp ~ continent, gapminder, sd)
```

Now for some plotting fabulousness:

```{r}
library(ggplot2)

```

```{r}
ggplot(gapminder, aes(x=continent, y=lifeExp)) +
  geom_boxplot(outlier.colour = "hotpink") +
  geom_jitter(position = position_jitter(width = 0.1, height = 0), alpha = 1/4)
```

Above is a plot of life expectancy by continent.

```{r}
ggplot(gapminder, aes(x=continent, y=gdpPercap)) +
  geom_boxplot(outlier.colour = "hotpink") +
  geom_jitter(position = position_jitter(width = 0.1, height = 0), alpha = 1/4)
```

Above is a plot of gross domestic product per capita.

```{r}
ggplot(gapminder, aes(x=continent, y=pop)) +
  geom_boxplot(outlier.colour = "hotpink") +
  geom_jitter(position = position_jitter(width = 0.1, height = 0), alpha = 1/4)
```

And above is a plot of population by continent.

```{r}
# integrating the country colors into the data frame
gap_with_colors <- data.frame(gapminder, cc=I(country_colors[match(gapminder$country, names(country_colors))]))

# now a bubble plot, focusing just on Africa and Europe in 2007

keepers <- with(gap_with_colors, continent %in% c("Africa", "Europe") & year == 2007)
plot(lifeExp~ gdpPercap, gap_with_colors,
    subset = keepers, log = "x", pch = 21,
    cex = sqrt(gap_with_colors$pop[keepers]/pi)/1500,
    bg = gap_with_colors$cc[keepers])
```


Above is a scatterplot of GDP per capita in 2007 with a dot representing each country in Asia and Europe. Presumable green is Europe and brown is Africa.

```{r}
# now a bubble plot, focusing just on the Americas and Asia in 1992

keepers <- with(gap_with_colors, continent %in% c("Americas", "Asia") & year == 1992)
plot(lifeExp~ gdpPercap, gap_with_colors,
    subset = keepers, log = "x", pch = 21,
    cex = sqrt(gap_with_colors$pop[keepers]/pi)/1500,
    bg = gap_with_colors$cc[keepers])
```



